Probability

### 2016-04-18 prob.tex
prob.tex \subsection{Plan} Add table...

\documentclass{article}
\usepackage{fullpage}
\usepackage{bbm}
\usepackage{bbold}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{multicol}

@@

\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}
\newcommand{\yutsung}[1]{\fbox{\begin{minipage}{0.9\textwidth}\color{purple}{Yu-Tsung says: #1}\end{minipage}}}
\newcommand{\ffzd}[1]{{\mathbb{F}^{d\;*}_{#1}}}

@@

Several assumptions are woven in the definition of a quantum probability space:
\begin{itemize}
\item the Hilbert space $\Hilb$;
\item the real interval $[0,1]$; 
\item the fact that each state induces a probability measure, i.e., the Born rule;
\item the fact that every probability measure is induced by a state,
  i.e., Gleason's theorem
\end{itemize}

In the remainder of the paper, we examine each of these assumptions
and consider variations motivated by computation in a world with limited
resources. In particular, we will consider a variant of the Hilbert
space over finite fields~$\mathbb{F}_{p^{2}}^{d}$. Instead of $\left[0,1\right]$,
we will consider set-valued probability measures, in particular $\left\{ 0\right\} $,
impossible, and $\left(0,\infty\right)$, possible. Surprisingly,
some combinations of space and probability will result in no probability
measure or an unique probability measure. In these cases, there is
no need to discuss whether there is a Born rule, because we do not
have enough probability to correspond to every state.

If there may be more than one probability measure, we will discuss
whether there is a Born rule to generate a probability measure from
a state. When the space is $\mathbb{C}^{d}$, we will try to induced
a Born rule from the conventional Born rule; when the space is $\mathbb{F}_{p^{2}}^{d}$,
there is no natural way to induce a probability measure from a state,
so we will set some conditions a Born~$\tilde{\pi}$ should have:
\begin{itemize}
\item Given a pure state $\ket{\Psi}\in\ffzd{p^{2}}$, a Born-rule~$\tilde{\pi}$
should give a probability~$\tilde{\pi}_{\Psi}$;
\item $\ip{\Psi}{\Phi}=0\Leftrightarrow\tilde{\pi}_{\Psi}\left(\ket{\Phi}\right)=\tilde{0}$,
where $\tilde{0}$ is $0$ while considering $\left[0,1\right]$ and
$\tilde{0}$ is impossible while considering $\left\{ \textrm{impossible},\textrm{possible}\right\} $.
\item $\tilde{\pi}_{\Psi}\left(\ket{\Phi}\right)=\tilde{\pi}_{\mathbf{U}\ket{\Psi}}\left(\mathbf{U}\ket{\Phi}\right)\textrm{ ,}$where
$\ket{\Psi},\ket{\Phi}\in\ffzd{p^{2}}$ and $\mathbf{U}$ is any unitary
map, i.e., $\mathbf{U}^{\dagger}\mathbf{U}=\mathbbm{1}$. 
\end{itemize}
Notice that when the space is $\mathbb{C}^{d}$, every Born rule we
consider will satisfy these three conditions.

Finally, if there is a Born rule, we will see whether every probability
measure is induced by a state, and establish Gleason's theorem.

>>>>

\documentclass{article}
\usepackage{fullpage}
\usepackage{bbm}
\usepackage{bbold}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{multicol}
\usepackage{array}
\usepackage{mathrsfs}

@@

\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}
\newcommand{\yutsung}[1]{\fbox{\begin{minipage}{0.9\textwidth}\color{purple}{Yu-Tsung says: #1}\end{minipage}}}
\newcommand{\ffzd}[1]{{\mathbb{F}^{d\;*}_{#1}}}
\def\C{{\mathbb{C}}}
\newcommand{\ff}[1]{\mathbb{F}_{#1}}

@@

Several assumptions are woven in the definition of a quantum probability space:
\begin{itemize}
\item the Hilbert space $\Hilb$;
\item the real interval $[0,1]$; 
\item the fact that each state induces a probability measure, i.e., the
Born rule~\cite{Born1984,Mermin2007}; 
\item the fact that every probability measure is induced by a state, i.e.,
Gleason's theorem~\cite{gleason1957,peres1995quantum,Redhead1987-REDINA}. 
\end{itemize}
In the remainder of the paper, we examine each of these assumptions
and consider variations motivated by computation in a world with limited
resources. In particular, we will consider a variant of the Hilbert
space over finite fields~$\mathbb{F}_{p^{2}}^{d}$~\cite{HansonOrtizSabryEtAl2015,DQT2014,geometry2013}.
Instead of $\left[0,1\right]$, we will consider set-valued probability
measures~\cite{Artstein1972,PuriRalescu1983}, in particular $\left\{ 0\right\} $,
impossible, and $\left(0,\infty\right)$, possible. Surprisingly,
some combinations of space and probability will result in no probability
measure or an unique probability measure. In these cases, there is
no need to discuss whether there is a Born rule, because we do not
have enough probability to correspond to every state.

If there may be more than one probability measure, we will discuss
whether there is a Born rule to generate a probability measure from
a state. When the space is $\mathbb{C}^{d}$, we will try to induced
a Born rule from the conventional Born rule; when the space is $\mathbb{F}_{p^{2}}^{d}$,
there is no natural way to induce a probability measure from a state,
so we will set some conditions a Born~$\tilde{\pi}$ should have: 
\begin{itemize}
\item Given a pure state $\ket{\Psi}\in\ffzd{p^{2}}$, a Born-rule~$\tilde{\pi}$
should give a probability~$\tilde{\pi}_{\Psi}$; 
\item $\ip{\Psi}{\Phi}=0\Leftrightarrow\tilde{\pi}_{\Psi}\left(\ket{\Phi}\right)=\tilde{0}$,
where $\tilde{0}$ is $0$ for $\left[0,1\right]$ and $\tilde{0}$
is impossible for $\mathscr{L}_{2}=\left\{ \textrm{impossible},\textrm{possible}\right\} $. 
\item $\tilde{\pi}_{\Psi}\left(\ket{\Phi}\right)=\tilde{\pi}_{\mathbf{U}\ket{\Psi}}\left(\mathbf{U}\ket{\Phi}\right)\textrm{ ,}$where
$\ket{\Psi},\ket{\Phi}\in\ffzd{p^{2}}$ and $\mathbf{U}$ is any unitary
map, i.e., $\mathbf{U}^{\dagger}\mathbf{U}=\mathbbm{1}$. 
\end{itemize}
Notice that when the space is $\mathbb{C}^{d}$, every Born rule we
consider will satisfy these three conditions.

Finally, if there is a Born rule, we will see whether every probability
measure is induced by a state, and establish Gleason's theorem. Notice
that Gleason's theorem only hold when $d\ge3$ in CQT so that we may
expect the situation for $d\ge3$ and $d=2$ are different.

The results can be summarized in the following table:\\
\begin{tabular}{>{\raggedright}m{0.15\columnwidth}>{\raggedright}m{0.15\columnwidth}>{\raggedright}m{0.15\columnwidth}>{\raggedright}m{0.15\columnwidth}>{\raggedright}m{0.15\columnwidth}>{\raggedright}m{0.15\columnwidth}}
\hline 
State space $\Hilb$ & Probability values & How many probability measure? & Is there a nature Born rule? & How many possible Born rule if there is no nature one? & How many probability measure not come from any possible Born rules?\tabularnewline
\hline 
\hline 
$\C^{2*}$ & $\left[0,1\right]$  & $\ge2$ & Yes &  & $\ge1$\tabularnewline
\hline 
$\C^{d*}$ for $d\ge3$ & $\left[0,1\right]$  & $\ge2$ & Yes &  & $0$\tabularnewline
\hline 
$\C^{d*}$ & $\mathscr{L}_{2}$ & $\ge2$ & Yes &  & $\ge1$\tabularnewline
\hline 
$\ff{3^{2}}^{d*}$ for $3\ge d\ge2$ & $\left[0,1\right]$  & $\ge2$ & No & $1$ & $\ge1$\tabularnewline
\hline 
$\ff{p^{2}}^{2*}$ for $p\ge7$ & $\left[0,1\right]$  & $\ge2$ & No & $\infty$ & $\ge1$\tabularnewline
\hline 
$\ff{7^{2}}^{3*}$ & $\left[0,1\right]$  & $1$ & No & $0$ & \tabularnewline
\hline 
$\ffzd{p^{2}}$for $d\ge3$ except $d=p=3$ & $\left[0,1\right]$  & $\ge1$ & No & $0$ & \tabularnewline
\hline 
$\ffzd{p^{2}}$ & $\mathscr{L}_{2}$ & $\ge2$ & No & $1$ & $\ge1$\tabularnewline
\hline 
\end{tabular}
